---
title: "session4"
author: "Pradeepti"
date: "12/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(dplyr)
require(tidyverse)
```

loading the dataset
```{r}
setwd("R:\\CT DPH\\R classes")
df = read_csv("2020-11-30TESTcases_scrub.csv")
```

exploring the dimensions of the dataset
```{r}
str(df)
summary(df)
```

Reduce dataset to first 10,000 cases
```{r}
newdf = df[1:10000, ]
str(newdf)
colnames(newdf)
```

Create a table of tests (tabulate the test types within the dataset)
Hint: Should have 
1 of COVID-19 PCR (CEPHEID) 
1 of SARS CoV 2 Ab (IgG) 
232 of SARS CoV 2 Ag (Quidel Sofia/Lumira), etcetera
```{r}
tests = newdf$test
head(tests)
```

There is one REALLY long string that is making your table hard to read (“SARS-CoV-2 (COVID-19) N gene [Cycle Threshold #] in Unspecified specimen by Nucleic acid amplification using CDC primer-probe set N1”). Replace this with “Unspecified Specimen” by over-writing those elements of the vector. Hint: use the grep function.
```{r}
bad_inputs = grep("Unspecified Specimen", tests)
head(bad_inputs)
tests[bad_inputs]="Unspecified specimen"
```

Re-tabulate your results again to confirm the change
```{r}
table(tests)
```

Lastly, notice that the virus name is formatted inconsistently in the test variable: either SARS CoV 2, or SARS-Cov-2 (dashes, or not). Using a new function, gsub, fix this.
```{r}
tests=gsub("SARS CoV 2","SARS-CoV-2",tests)
table(tests)
```


```{r}
sum(is.na(df$phone))
```


```{r}
```

