---
title: "Sessions 1-2-3"
output: html_notebook
---
loading the required packages
```{r}
require(tidyverse)
require(dplyr)
require(doBy)
```
R classes by Prof. Wininger from Yale.
Session #1
setting working directory and loading the dataset.
```{r}
setwd("R:\\CT DPH\\R classes")
data = read_csv("2020-11-30TESTcases_scrub.csv")
```

exploring the df, inspecting the dimensionality
```{r}
str(data)
head(data)
```


```{r}
summary(data)
dim(data)
nrow(data)
ncol(data)
glimpse(data)
```

inspecting the data types of variables in the dataset
```{r}
class(data$fname)
class(data$dob)
```

Session #2
subsetting the df for females age 20-40(inclusive) - answer Should be 20,846 females age 20-40
```{r}
subdata = subset(data, gender == 'Female' & age >= 20 & age <= 40)
subdata
```

take 2: using filter and piping
```{r}
data %>% filter(gender == 'Female' & age >= 20 & age <= 40)
```

checking if the subseting worked using table function
```{r}
table(data$age)
```

Show a proportional table of age via prop.table (HINT: nest your table command into the prop.table function!)
```{r}
prop.table(table(subdata$age))
```

Show a two-dimensional table of age versus ICU status (HINT: insert two arguments into your table function, and arrange age going left-to-right).
```{r}
table(subdata$icu,subdata$age)
```

Session #3
Subset syntax: subset(df name, subset OR select(condition for rows/columns))
summaryBy syntax: summaryBy(y~x, data = dfname, FUN = functionname)

reducing the dataset to first 2000 cases/observations
```{r}
data2000 = data[1:2000, ]
```

Reduce dataset to eventid, preg, gender, age
```{r}
#take1
#data2000 = data2000[ , c(1,20,8,7)]
#take2
data2000 = data2000[ , which(names(data2000) %in% c("eventid", "preg", "gender", "age"))]
```

Create a two-way table on preg versus gender - no pregnancies in males :) 
```{r}
table(data2000$gender, data2000$preg)
```

Summarize age per pregnancy status via summaryBy
take 1:using summaryBy
```{r}
summaryBy(age~preg, data = data2000, FUN = sum)

```

take 2: using piping and dplyr
```{r}
data2000 %>% group_by(preg) %>% summarize(Age = sum(age))
```


```{r}
```

